---
title: "Part 2 chi-squatr with lizards tails"
author: "Ali Martin"
date: "11/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)
library(broom)
```

## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species:

```{r}
lizards %>% 
  count(spp) %>% #how many times the species appears in the data set
  arrange(-n) #arrange in decreasing order
```

##Make a subset that contains:

-UTST lizards
-Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
-tail condition is recorded (W= whole, B = broken)

```{r}
utst <- lizards %>% 
  filter (spp =="UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c('B','W'))

```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor:: adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
```

Is there a significant association between site and tail breakage?

Do we think we have enough evidence to reject the null hypothesis which is there is no association between site and tail breakage

## Cpnvert to a contingency table

Use `column_to_rowname()` function.

```{r}
lizards_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi- square

```{r}
lizard_x2 <- chisq.test(lizards_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

Given a significance level of 0.05, Yes we have enough evidence to reject the null hypothesis of independence which means we would retain the alternative hypothesis that site and tail breakage are not independent

There is a significant association (ie non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic` , *p* = `r lizard_tidy$p.value`). 

Site and lizard tail breakage are not independent. 
